---
title: "Moodle-Quiz-Continuous-Conjugate"
author: "Jonas Barth"
date: "2023-04-25"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Task
Suppose we have observed the following 20 conditionally i.i.d. normal random variables with unknown mean $\theta$ (theta) and known variance $\sigma^2=100$.

```{r}
y = c( 103.25,88.37,88.95,104.81,92.27,75.01,80.79,83.2,99.9,91.26,91.67,88.77,82.1,105.53,92.64,67.99,94.67,82.93,76.98,85.48)
n = length(y)

known.var = 100
```

We want to carry out some conjugate Bayesian analysis using as a prior distribution a suitable distribution with mean $\mu_0 = 75$ and variance $\tau^2_0 = 1.25$.

```{r}
prior.mu = 75
prior.var = 1.25
```


# Provide the posterior mode as a point estimate
First, I need to update the $\mu$ and $\sigma$ parameters.

$$\begin{aligned}
\mu^* &= \frac{1}{\frac{1}{\tau^2_0} + \frac{n}{\sigma^2}} \left(\frac{1}{\tau^2_0} \mu_0 + \frac{n}{\sigma^2} \bar{y}\right) \\
\tau^2_* &= \left(\frac{1}{\tau^2_0} + \frac{n}{\sigma^2}\right)^{-1}
\end{aligned}$$

```{r}
post.mu = (((1/ prior.var) * prior.mu) + ((n / known.var) * mean(y)))
post.mu = post.mu / ((1 / prior.var) + (n / known.var))
post.var = 1 / ((1 / prior.var) + (n / known.var))
```

The mode of the posterior distribution should be $77.7657$ but is:
```{r}
post.mu
```

# Credible Interval 
Provide the shortest credible interval for the unknown parameter $\theta$ corresponding to 0.8 posterior probability:

```{r}
alpha = 0.8
library(TeachingDemos)

icdf = function(x) {
  return(qnorm(x, post.mu, sqrt(post.var)))
}

hpd(icdf, conf=alpha)
```
# Posterior Probability
Provide your posterior probability corresponding to the hypothesis that the unknown parameter $\theta \le 75.7657$.

```{r}
pnorm(75.7657, post.mu, sqrt(post.var))
```

# Prior Odds
What are the prior odds corresponding to the same hypothesis that the unknown parameter $\theta \le 75.7657$?

The prior odds are the probability of the event happening divided by the probability of the event **NOT** happening.
```{r}
prior.odds = pnorm(75.7657, prior.mu, sqrt(prior.var)) / (1 - pnorm(75.7657, prior.mu, sqrt(prior.var)))
prior.odds
```

# Bayes Factor
Compute the Bayes factor $B_{01}$ for comparing the two alternative hypotheses $H_0: \theta$  versus $H_1: \theta > 75.7657$.

To compute the Bayes Factor we divide the posterior odds by the prior odds.

```{r}
posterior.odds = pnorm(75.7657, post.mu, sqrt(post.var)) / (1 - pnorm(75.7657, post.mu, sqrt(post.var)))

bayes.factor = posterior.odds / prior.odds
bayes.factor
```





